---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
library(tidyverse)

# Read the data
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)

av %>% 
  select(
    Name.Alias,
    starts_with("Death")
  ) %>% 
  head()

deaths <- av %>% 
  pivot_longer(
    starts_with("Death"),
    names_to = "Time",
    values_to = "Died"
  ) %>%
  select(
    URL, Name.Alias, Time, Died
  )
head(deaths)

total_deaths <- deaths %>%
  filter(tolower(Died) == "yes") %>%
  nrow()

```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 


```{r}

average_deaths <- deaths %>%
  mutate(
    Time = parse_number(Time)
  ) %>% 
  group_by(URL, Died) %>% 
  summarise(
    total_death = max(Time)
  ) %>% 
  filter(Died != "") %>%
  ungroup() %>% 
  mutate(
    total_death = if_else(Died == "NO" & total_death == 1, 0, total_death) #Sets the people who never died to 0 deaths
  ) %>%
  summarise(
    average_total_death = mean(total_death, na.rm = TRUE)
  ) %>%
  pull(average_total_death) # Extract the value

print(average_deaths)

  
maxdeaths <- deaths %>% 
  mutate(
    Time = parse_number(Time)
  ) %>% 
  group_by(URL, Died) %>% 
  summarise(
    total_death = max(Time)
  ) %>%
  filter(Died != "")
  
maxdeaths %>% 
  ungroup() %>% 
  count(Died, total_death)

grouped_deaths <- maxdeaths %>%
  group_by(Died, total_death) %>%
  tally()
  
```

Similarly, deal with the returns of characters.

``` {r}

returns <- av %>% 
  pivot_longer(
    starts_with("Return"),
    names_to = "TimeReturned",
    values_to = "Returned"
  ) %>%
  select(
    URL, Name.Alias, TimeReturned, Returned,
  )

maxreturns <- returns %>% 
  mutate(
    TimeReturned = parse_number(TimeReturned)
  ) %>% 
  group_by(URL, Returned) %>% 
  summarise(
    total_return = max(TimeReturned)
  ) %>%
  filter(Returned != "")
  
maxreturns %>% 
  ungroup() %>% 
  count(Returned, total_return)

maxreturns %>%
  group_by(Returned, total_return) %>%
  tally()

```

Based on these datasets calculate the average number of deaths an Avenger suffers. 

The average death an avenger suffers is 0.5229, assuming we treat the Died=NO and total_death=1 as 0 deaths. However, if you just average the results from the maxdeaths (let the Died=NO and total_death=1 stay as 1 death) table you get 1.12069.

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.
Questions we can pick from 
2. "Jocasta — an android based on Janet van Dyne and built by Ultron — has been destroyed five times and then recovered five times."
3. "On 57 occasions the individual made a comeback."

Tirmidi Mohamed: Question: "I counted 89 total deaths."

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

Tirmidi's code -
```{r}
total_deaths <- deaths %>%
  filter(tolower(Died) == "yes") %>%
  nrow()

total_deaths
```

Megan's code - Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.
```{r}

at_least_once <- deaths %>%
  mutate(
    Time = parse_number(Time)
  ) %>% 
  group_by(URL, Died) %>% 
  summarise(
    total_death = max(Time)
  ) %>%
  filter(Died != "") %>%
  filter(Died == "YES") %>%
  nrow()
at_least_once

```

Next students code - 

```{r}

```
Next students code -

```{r}

```
### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.
**Tirmidi's answer**
 - The statement that there were a total of 89 deaths is true. I was able to check by tallying up all the yes for deaths and it adds up to 89 so this is true.
 
**Megan's answer**
 - The print statement shows that the number of avengers who have died at least once is equal to 69. I used the deaths table, calculated the total deaths similar to the maxdeaths table, and filtered if they had died or not. Each row represents a unique avenger who has died, so the count of rows, which is 69, proves that 69 avengers died at least once.

Upload your changes to the repository. Discuss and refine answers as a team.
